---
title: "Covid Trial Design Modification Simulations (C163)"
author: "ADCS Biostatistics Core"
date: "`r Sys.Date()`"
fontsize: 11pt
output:
  pdf_document:
    toc: true
    number_sections: true
    highlight: tango
    keep_tex: true 
    includes:
      in_header: "/Users/zenn/prj/ADCSpreamble.tex"
header-includes:
  - \rhead{C163 project. COVID Design report (draft)}
bibliography: /Users/zenn/shr/library.bib
latex_engine: xelatex
---

\lhead{COVID Design Modifications Analysis}
\rhead{\thepage} 
\listoffigures
\listoftables

```{r  include=FALSE}
options(dplyr.summarise.inform = FALSE, dplyr.print_max = 1e9)
library(pacman)
p_load(purrr, ztable, patchwork,ggthemes,  
  kableExtra, showtext, skimr, broom.mixed, lme4, lsmeans, lubridate, rsample,
  tidyverse, knitr, broom, nlme, purrr, tidyr
)

pd <- position_dodge(.20)
opts_chunk$set(
  warning = F, message = F, echo = F, fig.width = 4.2,
  fig.height = 4, results = "asis", dev = "pdf",
  opts_chunk$set(cache.path = "cache/")
)
source("~/shr/zz.tools.R")
source("c163_fns.R")

options(scipen = 1, digits = 2, warn = -1, ztable.type = "latex")
showtext_auto()
# aa = font_files()    
# grep(aa$family, "School")
#syntax:font_add(family ="<family_name>",regular="/pth/to/font/file")
font_add("Montserrat", "Montserrat-Regular.otf")
font_add("Montserrat Thin", "Montserrat-Thin.otf")
font_families()
myfont1 = "Montserrat"
myfont2 = "Montserrat Thin"
```


\input{"/Users/zenn/shr/zz.tools.tex"}


# Introduction

\input{intro.txt}

```{r cache=F}
seed = 47
set.seed(seed)
```

# Methods

Analyses were based on ADAS 11 data from 641 patients/completers from the
pooled data sets for the ADCS Homocystine and Lipid-Lowering trials. 

Given target sample size, say N, and effect size (mean change score
difference at 12 months), say delta, and
pooled standard deviation, s,  individual profiles were selected at
random from the pool per the following algorithm: 

1. least squares slopes were computed for each profile. 

2. the set of 641 slopes were ordered from smallest to largest. 

3. A set, N/2, of "placebo" group slopes was selected at random
without replacement using sampling weights biased toward larger
slopes, (see figure 1). 

4. A second set of N/2 "active" slopes were selected in a similar
manner using sampling weights biased toward smaller slopes. 


5. A dropout process was applied to the resulting data set of size N;
30\% of subjects were randomly selected and 50\% of those were assumed
to drop after 6 months of followup and 50\% before 6 months of followup. 

6. A COVID19 pause process was applied, one of three, (see figure 2),
filtering the dataset per the anticipated loss (or delay) of subject
visit information. 

7. For each combined and filtered (both dropout and pause scenario)
sample calculate the observed delta at 12 months, and the pooled
standard deviation, s. and compare to defined tolerance (specified at
.05 for delta and .2 for s)  

8. Repeat steps 3 through 7 until delta and s are within tolerance. 

7. Repeat steps 3 through 8 until four data sets matching criteria are
determined. 

9.  Repeat overall process for multiple values of N and delta. 7. For
this experiment assume a dropout process impacting 30\% of subjects
and a fixed target s of 6.0.

For each of the parameter profiles and under three different COVID19
trial pause scenarios, matching data sets were in turn submitted
to a bootstrap analysis with 1000 replicates to determine the power of
three statistical testing procedures, .   

The bootstrap process entails resampling the dataset multiple times
without replacement, calculating the three test statistics and their
associated p-values. The proportion of p-values less than 0.05 over the 1000
replicates is an estimate of the power of the test. 



```{r  eval=T, cache=F}
source("~/prj/c163/c163_preprocess.R")
dat1 <- dat1 %>%
  fil(!trial == "dha") %>%
  sel(-arm, -trial)

```

```{r  eval=T, cache=T, child="c163_viz.Rmd"}
```


```{r include=F, cache=F}

output_dir <- file.path(".", "data")
if (!dir.exists(output_dir)){
dir.create(output_dir) } 

tdelta <- c(2.0, 2.6, 3.0, 4.0)
tsize <- c(320, 360, 400)
filtervar <- c("filter0","filter1", "filter2", "filter3")
vers <- 1:20

params <- expand.grid(tsize, tdelta, vers, filtervar) %>%
  setNames(c("tsize", "tdelta", "vers", "filtervar")) %>%
  mutate(filtervar = as.character(filtervar)) 

ff <- function(x) {
  pmap(x, function(tsize, tdelta, vers, filtervar) {
    gendata(tsize, tdelta, vers, filtervar)
  })
}
#gendata(tsize=320, tdelta=3.0, vers=5, filtervar="filter0")
#gendata(tsize=360, tdelta=4.0, vers=5, filtervar="filter0")
#gendata(tsize=400, tdelta=2.0, vers=5, filtervar="filter0")


#params %>% ff

sims <- data.frame(
  test = character(0), power = numeric(0), tsize = numeric(0),
  delta = numeric(0), reps = numeric(0), scenario = numeric(0), vers = numeric(0)
)

output_dir <- file.path(".", "sim_results")
if (!dir.exists(output_dir)){
dir.create(output_dir)
} 

#write_csv(sims, "sim_results/sim.csv", append = F)

reps <- 10

fnames <- list.files("./data", "^dat_n") 
fnames2 <- fnames %>%
  map_dfr(., function(x) {
    strsplit(x, "\\D") %>%
      unlist() %>%
      setNames(letters[1:15])
  }) %>%
  sel(tsize = f, tdelta = h, scenario = j, vers = l) %>%
  cbind(fname = fnames, reps = reps, .)

ff <- function(x) {
  pmap(x, function(fname, reps, tsize, tdelta, scenario, vers) {
    boots(fname, reps, tsize, tdelta, scenario, vers)
  })
}


fnames2 = slice(fnames2,c(6,12,14))


fnames2 %>% ff

results <- read.csv("sim_results/sim.csv")
results <- read.csv("sim_results/sim_archive/sim_2020-08-06-0854.csv")
#results1 <- read.csv("sim_results/sim_2020-08-03-1810.csv")
#results2 <- read.csv("sim_results/sim_2020-07-31-1646.csv")
#results3 <- read.csv("sim_results/sim_2020-08-01-0944.csv")
#results4 <- read.csv("sim_results/sim_2020-08-04-0844.csv")

#resultssum <- rbind(results,results1, results2, results3,  results4) %>%

resultssum <- rbind(results) %>%
	#   fil(!test == "ztest")  %>%  
	#   fil(delta %in% c(20,26,30,40))  %>%  
  group_by(test, tsize, delta, scenario) %>%
  summ(n = n(), treps = sum(reps), pow = median(power), se=sd(power)/(n^.5)) %>%
  mut(
    ci_low = pow - 1.96 * ((pow * (1 - pow) / treps)^.5),
    ci_low2 = pow - 1.96 * se,
    ci_upper = pow + 1.96 * ((pow * (1 - pow) / treps)^.5),
    ci_upper2 = pow + 1.96 * se  ) %>%
  mut(
    scenario = as.character(scenario),
    test = recode(test, "lme_cat" = "lme-cat", "lme_slp" = "lme-slp")
  )
   
```

\newpage

# Simulation Results



```{r eval=T, cache=F, fig.width=6.2, fig.height=8 }

options(digits = 6)
resultssum %>%
  mut(delta = delta / 10, conf_int = paste0("(", round(ci_low, 3), ", ", round(ci_upper, 3), ")")) %>%
  sel(test, scenario, tsize, delta, pow, conf_int) %>%
  plyr::arrange(test, scenario, tsize, delta) %>%
  setNames(gsub("_", "-", names(.))) %>%
  kable("latex", digits = 5, escape = F, booktab = T, longtable = T) %>%
  kable_styling(latex_options = "striped") %>%
  row_spec(0, angle = -60) %>%
  row_spec(c(6, 12), hline_after = T)
# 
#  %>%
#   column_spec(5, bold = T) 
cat("\n \\captionof{table}{Simulation results}\\vspace{1in}\n")

cbPalette <- c("#000000","#9900bb","#35bb00", "#0061bb", "#bb0000")
#gg = ggplot(data = resultssum, aes(x = delta / 10, y = pow, group=test)) +
  #geom_point(shape = 1, alpha=.5) +
 #theme_bw(base_family = myfont1) +
  #xlim(1, 5) +
  #facet_grid(scenario ~ tsize, labeller = label_both) +
  #geom_errorbar(aes(ymin = ci_low, ymax = ci_upper), width = 0.2, color = "black", alpha = .3) +
  #geom_hline(yintercept = 0.8, color = "red", alpha = .5, size = .3) +
  #geom_vline(xintercept = 2.6, color = "red", linetype = 6, alpha = .5, size = .3) +
  #xlab("Delta at 12 months, completers") +
  #theme(strip.text = element_text(size = 5)) +
  #theme(axis.text.y = element_text(size = 5)) +
  #scale_color_manual(values=cbPalette)+
  #scale_y_continuous(breaks = seq(0, 1, .2))+ 
     #geom_smooth(method = "glm", method.args = list(family = binomial(link = "probit") ), 
                 #aes(weight = 1000, color = "probit smooth"), se = F, size=.5 ) 
 #ggsave("delta.pdf")

tsize <- c(320,360, 400)
delta <- c(2, 2.6, 3, 4)
powers <- expand.grid( tsize,  delta) %>% setNames(c( "tsize",  "delta"))
ff <- function(x) {
  pmap_dbl(x, function( tsize, delta) {
    power.t.test(n = .7 * tsize / 2, delta = delta, sd = 6, sig.level = .01)$power
  }) }

powers <- powers %>%
  mut(pow = powers %>% ff()) %>%
  mut(delta = as.numeric(delta)) %>% mut(tsize=as.integer(tsize))  %>% mut( 
scenario = "orig", test="orig", delta = 10*delta)

powers1 = powers %>%  mut(test="lme-cat")
powers2 = powers %>%  mut(test="lme-slp")
powers3 = powers %>%  mut(scenario="1")
powers4 = powers %>%  mut(scenario="2")
powers5 = powers %>%  mut(scenario="3")
powers6 = powers %>%  mut(scenario="0")
resultssum2 = resultssum %>%  full_join(powers1) %>% full_join(powers2) 
resultssum3 = resultssum %>%  full_join(powers3) %>% full_join(powers4)  %>%  
full_join(powers5) %>% full_join(powers6) 


write.csv(resultssum3, "symptomatic_pow.csv")




cbPalette <- c("#35bb00", "#0061bb", "#bb0000")
gg <- ggplot(data = resultssum3, aes(x = delta / 10, y = pow, group = test, color = test)) +
 geom_point(alpha = .5) +
 facet_grid(scenario ~ tsize, labeller = label_both) +
 xlim(1, 5) +
 geom_hline(yintercept = 0.8, color = "red", alpha = .5, size = .3) +
 geom_vline(xintercept = 2.6, color = "red", linetype = 6, alpha = .5, size = .3) +
 xlab("Effect Size") +
 theme_bw(base_family = myfont1) +
 theme(strip.text = element_text(size = 5)) +
 theme(axis.text.y = element_text(size = 5)) 
# +
#   scale_color_manual(values=cbPalette)+
#  scale_y_continuous(breaks = seq(0, 1, .2))
# +
 #      geom_smooth(method = "glm", method.args = list(family = binomial(link = "probit") ), 
 #                  aes(weight = 1000, color = test), se = F, size=.5) 
 ggsave("deltatest.pdf")

cbPalette <- c("#000000","#9900bb","#35bb00", "#0061bb", "#bb0000")
gg  <- ggplot(data = resultssum2, aes(x = delta / 10, y = pow, group = scenario, color = scenario)) +
geom_point(alpha=.2, position=pd) +
geom_errorbar(aes(ymin=ci_low2, ymax=ci_upper2),width = 0.1, alpha=.3, position=pd)+
 facet_grid(test ~ tsize, labeller = label_both) +
 xlim(1, 5) +
 geom_hline(yintercept = 0.8, color = "red", alpha = .5, size = .3) +
 geom_vline(xintercept = 2.6, color = "red", linetype = 6, alpha = .5, size = .3) +
 xlab("Effect Size") +
 theme_bw(base_family = myfont1) +
 theme(strip.text = element_text(size = 5)) +
 theme(axis.text.y = element_text(size = 5)) +
  scale_color_manual(values=cbPalette)+
 scale_y_continuous(breaks = seq(0, 1, .2))+
 geom_smooth(method = "glm", method.args = list(family = binomial(link = "probit") ), 
                 aes(weight = 1000, color = scenario), se = F, size=.5, position=pd) 
 ggsave("deltascenario.pdf")

dat4 = resultssum3
dat4 = dat4  %>%  sel(Power =pow, N = tsize, everything()) %>% mut(effect = delta/10, test = recode(test, "ztest" = "ttest")) 
yy = ggplot(data = dat4, aes(x = effect, y = Power, group = test, color = test)) +
  geom_point(shape = 16, size = 1.15) +
  facet_grid(scenario ~ N, labeller = label_both) +
  theme_bw() +
  geom_hline(yintercept = 0.8, color = "red", alpha = .5, size = .3) +
  geom_vline(xintercept = 2.6, color = "red", linetype = 6, alpha = .5, size = .3) +
  xlab("aDAS-cog") +
  theme(strip.text = element_text(size = 10)) +
  theme(axis.text.y = element_text(size = 10)) +
  scale_y_continuous(breaks = seq(0, 1, .2)) +
  scale_x_continuous(limits=c(1.5, 3),breaks = seq(1.0, 3, .5)) +
  scale_color_manual(values=c('red','purple','blue','green')) +
  stat_smooth(method="glm", method.args=list(family=binomial(link="probit")), se=F, size = 0.75)

 ggsave("yy3.pdf", width=9, height=12)

```

\includegraphics{deltatest.pdf}
\captionof{figure}{Simulation results - by sample size and scenario - 
two mmrm tests per facet: lme-cat (categor. time), lme-slp (cont. time), 
also original t-test calc added to each plot}\vspace{1in}
\includegraphics{deltascenario.pdf}
\captionof{figure}{Simulation results. Smoothed inverse probit curves.
bootstrap point estimates and 95\% confidence intervals - by sample size and test - four
scenarios per facet, also for reference original t-test calc 'orig' added to each plot. }\vspace{1in}
\newpage
```

# Appendix

```{r eval=T, child="demog.Rmd"}
```
<!--```{r eval=T, child="appendix.Rmd"}-->
<!--```-->


\appendix

\section{Full-size images}

\unvbox\savedimgs
